Preventing a Nuclear ISIL: \\Recommendations for United States Policy

### 2015-05-08 draft.tex
Broke out Chapter 4 (Nuclear Physics Primer) into 4 Appendices.

\title{Our Title}

\date{Some date}

@@

This section is meant to serve as a basic guide to technical concepts necessary for understanding nuclear weapons. Although the content has been distilled to the most fundamental information and is deliberately simplified, it is critical to understanding the scientific context of nuclear weapons \cite{Prussin2014}. For flow of this report, the Nuclear Weapons Primer has been broken into 4 appendices:

\section{???????  Call out cross ref'd appendix names here ?????????}

@@

\section{??????? Move this section into its own appendix, following the glossary appendix ???????? }

Appendix XXXXXXX  covers the basic nuclear physics needed to understand the functioning of a a nuclear weapon. It covers topics such as nuclear structure, criticality, fission, fusion, and energy release, all of which are key to framing the development of a nuclear weapon. Additionally, it covers radioactive decay, radiation interactions with matter, and biological effects of radiation, useful for understanding the consequences of successful detonation of nuclear weapons or RDDs. These concepts and definitions will be used throughout the report to explain and formulate technical and policy solutions to eliminate pathways for ISIL acquisition of a nuclear weapon.

\subsection{Nuclear Structure}

\subsection{Fission}

\subsection{Criticality}

\subsection{Critical Mass}

\subsection{Fusion}

\subsection{Energy Release}

\subsection{Radioactive Decay}

\subsection{Radiation Interactions with Matter}

\subsection{Health Physics}

@@

\section{??????? Move this section into its own appendix, following the glossary appendix ???????? }

Appendix XXXXXXX  covers the basics of a nuclear fuel cycle necessary for both civilian and military uses of nuclear energy. The focus is on technologies and capabilities required to develop an indigenous nuclear weapons program. The Appendix addresses both the plutonium and uranium fuel cycles, and compares differences between the two. Any pathway to obtaining a nuclear weapon (other than theft or purchase of a complete nuclear weapon) will require one or more of the capabilities outlined in Appendix XXXXXXX.

\subsection{Uranium}

\subsection{Plutonium}

@@

\section{??????? Move this section into its own appendix, following the glossary appendix ???????? }

Appendix XXXXXXX  covers the basic categories of weapons enabled by nuclear materials: RDDs, fission weapons, and fusion weapons. This Appendix builds on the concepts outlined in

\section{??????? Cross ref to last two appendices ???????}

. Any ISIL pursuit of nuclear weapons or nuclear technology would ultimately result in the development of one of these three types of weapons.

\subsection{RDD} \label{sec:RDD}

\subsubsection{RDD Case Study}  \label{sec:case_study}

\subsubsection{Radioisotopes of Concern}

\subsection{Fission}

\subsection{Fusion}

@@

@@

\section{Purchase or Gift}

@@

Lead Agency: Department of Defense

Contributing Agency: Intelligence Community \normalfont

@@

While it is extremely unlikely ISIL would attempt an indigenous nuclear program for the reasons detailed in \autoref{sec:fuel_cycle}, if ISIL attempted to develop an indigenous program, airstrikes could easily derail the program. As detailed in \autoref{sec:fuel_cycle}, the fuel for a nuclear weapon requires either large enrichment facilities or a nuclear reactor, depending on the type of fuel used. In either case, the necessary facilities are very large and would be easily detectable by Intelligence, Surveillance, and Reconnaissance operations focusing on Iraq and Syria.

@@

\section{If ISIL detonates a nuclear device}

If ISIL is successful in detonating a nuclear device, the U.S. must be prepared to take military action in response, as well as to prevent the detonation of a second device. Depending on the U.S. assessment of the threat posed by ISIL's attack (which is a discussion that must be had at the highest levels), a military response may range from an increased air campaign to U.S. ground force involvement in a campaign to destroy ISIL. Additionally, the U.S. must be prepared to collect forensic data to determine, if possible, if any additional countries were complicit in ISIL acquiring a nuclear device. If another nation aided ISIL, the U.S. should consider military action in response consistent with the recommended declaratory policy in

Section XXX Forensics Policy Section XXX.

\chapter{Dissention Minority Opinion (Or Option B or Some other name)}

@@

RPMs evolved out of a layered defense concept to prevent a low-probability/high-impact scenario, nuclear terrorism. This layered defense is composed of RPMs as discussed in

@@

Section XXX INSERT REF TO RPM section here XXX, the overseas equivalent known as the SLD as discussed in

Section XXX INSERT REF TO SLD SECTION HERE XXX, and the Global Threat Reduction Initiative discussed in

@@

The sum result of this discussion is that we are at best facing the prospect of a nuclear armed ISIL in several years assuming no disruptions due to discovery and outside assistance from a nuclear capable state, both of which are counter to the history of every nuclear program \cite{pollack2014unthinkable,Reed2010}.  Given this fact, the best policy we have is to place financial, diplomatic, and military pressure on ISIL, all of which are current policies of the Obama Administration as shown in

Table XXX POLICY CHART XXX.  Additionally, there are no identified potential proliferation pathways that are not covered to some extent by current policies, programs, or operations.  As war was building in Europe in 1940, Edward Murrow reported, \enquote{The people here feel the machine is out of control, that we are all passengers on an express train traveling at high speed through a dark tunnel toward an unknown destiny.  The suspicion recurs that the train may have no engineer.}  The proposed policies are emblematic of a machine out of control in support of Dick Cheney's \enquote{1\% Doctrine.} Therefore, the dissenting opinion recognizes that the marginal cost to implement the proposed policies outweigh the marginal benefit and recommends to not repeat the mistakes of history by implementing the proposed policies of this report. Instead, it is proposed that the U.S. implements no changes to U.S. policy specifically to prevent the low probability event that is ISIL obtaining nuclear weapons. This would entail \enquote{staying the course} with President Obama's current policy of disrupting, degrading, and ultimately destroying ISIL as a necessary and sufficient condition to prevent nuclear weapon acquisition, while not losing focus on the real ISIL national security challenges \cite{WhiteHouse2014}.

@@

\chapter{???????   Budget, cohesive implemented plan   ?????????}

@@

% \appendix
\chapter[Appendix B: Intelligence Overview]{Intelligence Overview}  \label{app:intel}

\chapter[Appendix C: International Partnerships]{International Partnerships} \label{app:partners}

\chapter[Appendix D: Technology]{Technology} \label{app:tech}

@@

\chapter[Appendix E: ????????]{?????????} \label{app:relations_table}

>>>>

\title{Preventing a Nuclear ISIL: \\Recommendations for United States Policy}

\date{11 May 2015}

@@

This section is meant to serve as a basic guide to technical concepts necessary for understanding nuclear weapons. Although the content has been distilled to the most fundamental information and is deliberately simplified, it is critical to understanding the scientific context of nuclear weapons \cite{Prussin2014}. For flow of this report, the Nuclear Weapons Primer has been broken into 4 appendices: \autoref{app:glossary}, \autoref{app:physics}, \autoref{app:app_fuel_cycle}, and \autoref{app:weapons}.

@@

\autoref{app:physics}  covers the basic nuclear physics needed to understand the functioning of a a nuclear weapon. It covers topics such as nuclear structure, criticality, fission, fusion, and energy release, all of which are key to framing the development of a nuclear weapon. Additionally, it covers radioactive decay, radiation interactions with matter, and biological effects of radiation, useful for understanding the consequences of successful detonation of nuclear weapons or RDDs. These concepts and definitions will be used throughout the report to explain and formulate technical and policy solutions to eliminate pathways for ISIL acquisition of a nuclear weapon.

@@

\autoref{app:app_fuel_cycle}  covers the basics of a nuclear fuel cycle necessary for both civilian and military uses of nuclear energy. The focus is on technologies and capabilities required to develop an indigenous nuclear weapons program. The Appendix addresses both the plutonium and uranium fuel cycles, and compares differences between the two. Any pathway to obtaining a nuclear weapon (other than theft or purchase of a complete nuclear weapon) will require one or more of the capabilities outlined in \autoref{app:app_fuel_cycle}.

@@

\autoref{app:weapons}  covers the basic categories of weapons enabled by nuclear materials: RDDs, fission weapons, and fusion weapons. This Appendix builds on the concepts outlined in \autoref{app:physics} and \autoref{app:app_fuel_cycle}. Any ISIL pursuit of nuclear weapons or nuclear technology would ultimately result in the development of one of these three types of weapons.

@@

\subsection{??????? Shrink down pictures, two to a page ????????}

@@

\section{Purchase or Gift} \label{sec:purchase_gift}

@@

@@

While it is extremely unlikely ISIL would attempt an indigenous nuclear program for the reasons detailed in \autoref{app:app_fuel_cycle}, if ISIL attempted to develop an indigenous program, airstrikes could easily derail the program. As detailed in \autoref{app:app_fuel_cycle}, the fuel for a nuclear weapon requires either large enrichment facilities or a nuclear reactor, depending on the type of fuel used. In either case, the necessary facilities are very large and would be easily detectable by Intelligence, Surveillance, and Reconnaissance operations focusing on Iraq and Syria.

@@

\section{Post-Detonation U.S. Response}

If ISIL is successful in detonating a nuclear device, the U.S. must be prepared to take military action in response, as well as to prevent the detonation of a second device. Depending on the U.S. assessment of the threat posed by ISIL's attack (which is a discussion that must be had at the highest levels), a military response may range from an increased air campaign to U.S. ground force involvement in a campaign to destroy ISIL. Additionally, the U.S. must be prepared to collect forensic data to determine, if possible, if any additional countries were complicit in ISIL acquiring a nuclear device. If another nation aided ISIL, the U.S. should consider military action in response consistent with the recommended declaratory policy in \autoref{sec:purchase_gift}.

\chapter{Dissenting Opinion}

@@

RPMs evolved out of a layered defense concept to prevent a low-probability/high-impact scenario, nuclear terrorism. This layered defense is composed of RPMs as discussed in \autoref{sec:domestic_protection}, the overseas equivalent known as the SLD as discussed in \autoref{sec:Trafficking}, and the Global Threat Reduction Initiative discussed in

@@

6.6. something

@@

The sum result of this discussion is that we are at best facing the prospect of a nuclear armed ISIL in several years assuming no disruptions due to discovery and outside assistance from a nuclear capable state, both of which are counter to the history of every nuclear program \cite{pollack2014unthinkable,Reed2010}.  Given this fact, the best policy we have is to place financial, diplomatic, and military pressure on ISIL, all of which are current policies of the Obama Administration as shown in \autoref{fig:policy_chart}.  Additionally, there are no identified potential proliferation pathways that are not covered to some extent by current policies, programs, or operations.  As war was building in Europe in 1940, Edward Murrow reported, \enquote{The people here feel the machine is out of control, that we are all passengers on an express train traveling at high speed through a dark tunnel toward an unknown destiny.  The suspicion recurs that the train may have no engineer.}  The proposed policies are emblematic of a machine out of control in support of Dick Cheney's \enquote{1\% Doctrine.} Therefore, the dissenting opinion recognizes that the marginal cost to implement the proposed policies outweigh the marginal benefit and recommends to not repeat the mistakes of history by implementing the proposed policies of this report. Instead, it is proposed that the U.S. implements no changes to U.S. policy specifically to prevent the low probability event that is ISIL obtaining nuclear weapons. This would entail \enquote{staying the course} with President Obama's current policy of disrupting, degrading, and ultimately destroying ISIL as a necessary and sufficient condition to prevent nuclear weapon acquisition, while not losing focus on the real ISIL national security challenges \cite{WhiteHouse2014}.

@@

@@

\chapter[Appendix B: Basic Nuclear Physics \& Definitions]{Basic Nuclear Physics \& Definitions}  \label{app:physics}

\section{Nuclear Structure}

\section{Fission}

\section{Criticality}

\section{Critical Mass}

\section{Fusion}

\section{Energy Release}

\section{Radioactive Decay}

\section{Radiation Interactions with Matter}

\section{Health Physics}

\chapter[Appendix C: Fuel Cycle]{Fuel Cycle}  \label{app:app_fuel_cycle}

\section{Uranium}

\section{Plutonium}

\chapter[Appendix D: Weapons of Interest]{Weapons of Interest}  \label{app:weapons}

\section{RDD} \label{sec:RDD}

\subsection{RDD Case Study}  \label{sec:case_study}

\subsection{Radioisotopes of Concern}

\section{Fission}

\section{Fusion}

\chapter[Appendix E: Technology]{Technology} \label{app:tech}

@@

% \appendix
\chapter[Appendix F: Intelligence Overview]{Intelligence Overview}  \label{app:intel}

\chapter[Appendix G: International Partnerships]{International Partnerships} \label{app:partners}

\chapter[Appendix H: ????????]{?????????} \label{app:relations_table}
